---
title: "Project 2"
author: "Nicole Levin, Shaoyu Wang"
date: "2022-10-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This document is a vignette that describes pulling data from an API and performing some exploratory data analysis based on those results. This vignette will utilize the Spoonacular food API. More information about this API can be found [here](https://spoonacular.com/food-api) In this vignette, we will be building some functions to interact with the API and doing some exploratory data analysis based on some example search results.  

# Required Packages
The following packages are required for interacting with the functions in this vignette:
- httr
- dplyr
- jsonlite
- tidyverse

First, we will load the required packages
```{r message=FALSE, warning=FALSE}
#Load libraries
library(httr)
library(dplyr)
library(jsonlite)
library(tidyverse)
```

# Building Search Functions
There are three functions that have been created to allow the user to make a search from the API. 

## `ingredient_search`

The first is called `ingredient_search` and it allows the user to select an ingredient and a number of results to return. The function will then provide an output data frame with many points of information for each of the recipes. The function works by first searching for recipes with the ingredient. The ID numbers that the API identifies these recipes by are then plugged into a second search to get a combination of quantitative and categorical data including health scores, aggregate likes, whether a dish is vegetarian, whether it is gluten free, and whether it is dairy free. Full search code is as follows:  

```{r}
#User options for search. Can input: (1)ingredient for search, (2)Number of results
ingredient_search <- function(ingredient, number=10){
  ingredient <- tolower(ingredient)
  search_url <- "https://spoonacular-recipe-food-nutrition-v1.p.rapidapi.com/recipes/findByIngredients"
  queryString <- list(
  ingredients = ingredient,
  number = number)
search_result <- VERB("GET", search_url, add_headers('X-RapidAPI-Key' = '73a2057416msh633772bf8275edfp1b3775jsn7e44cc5cb98d', 'X-RapidAPI-Host' = 'spoonacular-recipe-food-nutrition-v1.p.rapidapi.com'), query = queryString, content_type("application/octet-stream"))
  parsed_search_result <- fromJSON(rawToChar(search_result$content))
  id_list <- paste(as.character(parsed_search_result$id), collapse=",")
bulk_url <- "https://spoonacular-recipe-food-nutrition-v1.p.rapidapi.com/recipes/informationBulk"
queryString <- list(ids = id_list)
bulk_search <- VERB("GET", bulk_url, add_headers('X-RapidAPI-Key' = '73a2057416msh633772bf8275edfp1b3775jsn7e44cc5cb98d', 'X-RapidAPI-Host' = 'spoonacular-recipe-food-nutrition-v1.p.rapidapi.com'), query = queryString, content_type("application/octet-stream"))
  parsed_bulk_search <- fromJSON(rawToChar(bulk_search$content))
  results_df <- as_tibble(parsed_bulk_search) %>% select(id, title, glutenFree, dairyFree, vegetarian, aggregateLikes, healthScore, weightWatcherSmartPoints, sourceName, pricePerServing, readyInMinutes)
  combined_results <- results_df %>% mutate(WW_category = if_else(weightWatcherSmartPoints <10, 1, if_else(weightWatcherSmartPoints <20, 2, 3))) 
  factor(combined_results$WW_category, levels = c(1, 2, 3), labels = c("low", "medium", "high"))
return(combined_results)
}
#Run a test
ingredient_test <- ingredient_search(ingredient="cherry", number=50)
ingredient_test
```

The example search results are for the ingredient "cherry with a maximum number of results of 30. As the output shows, we can get a lot of information from a simple search.  

## `calorie_search`
The second function is called "calorie_search" and it allows the user to select a minimum number of calories, maximum number of calories, whether they want random results or ordered, and what number of results they want. This function has default values for all variables, so they do not all have to be specified for the search to execute. It works in a similar fashion, where the initial query gets the list of ID numbers to be used in a second query to get more data. In this case though, the initial nutrition search based on calories gives us some variables like number of calories, grams of carbohydrates, grams of protein, and grams of fat that we can bring into the final data set too. The full code of the function is as follows:  

```{r}
#This function is a search by calories. User options are (1)Minimum calories, (2)Maximum calories, (3)Whether we want a random set of results or the first ones alphabetically, and a number of results. All of these have defaults, so the user can specify nothing and still get a result.
calorie_search <- function(min_calories=0, max_calories=5000, random=FALSE, number=10){
  min_calories <- as.character(min_calories)
  max_calories <- as.character(max_calories)
  number <- as.character(number)
  search_url <- "https://spoonacular-recipe-food-nutrition-v1.p.rapidapi.com/recipes/findByNutrients"
  if(random==FALSE){
    queryString <- list(minCalories = min_calories, maxCalories = max_calories, number=number)}
  else {queryString <- list(minCalories = min_calories, maxCalories = max_calories, random = TRUE)}
  search_result <- VERB("GET", search_url, add_headers('X-RapidAPI-Key' = '73a2057416msh633772bf8275edfp1b3775jsn7e44cc5cb98d', 'X-RapidAPI-Host' = 'spoonacular-recipe-food-nutrition-v1.p.rapidapi.com'), query = queryString, content_type("application/octet-stream"))
  parsed_search_result <- fromJSON(rawToChar(search_result$content))
  id_list <- paste(as.character(parsed_search_result$id), collapse=",")
  bulk_url <- "https://spoonacular-recipe-food-nutrition-v1.p.rapidapi.com/recipes/informationBulk"
queryString <- list(ids = id_list)
bulk_search <- VERB("GET", bulk_url, add_headers('X-RapidAPI-Key' = '73a2057416msh633772bf8275edfp1b3775jsn7e44cc5cb98d', 'X-RapidAPI-Host' = 'spoonacular-recipe-food-nutrition-v1.p.rapidapi.com'), query = queryString, content_type("application/octet-stream"))
  parsed_bulk_search <- fromJSON(rawToChar(bulk_search$content))
  results_df <- as_tibble(parsed_bulk_search) %>% select(id, title, glutenFree, dairyFree, vegetarian, aggregateLikes, healthScore, weightWatcherSmartPoints, sourceName, pricePerServing, readyInMinutes)
  results_df <- results_df %>% mutate(WW_category = if_else(weightWatcherSmartPoints <10, 1, if_else(weightWatcherSmartPoints <20, 2, 3))) 
  factor(results_df$WW_category, levels = c(1, 2, 3), labels = c("low", "medium", "high"))
  combined_results <- as_tibble(c(results_df, parsed_search_result[,3:8]))
  return(combined_results)
}
calorie_test <- calorie_search(min_calories = 100, max_calories = 500, number=50)
calorie_test
```

Both of these two search function above included an output variable of "WW_category" This variable did not exist in the source data and had to be created using the Weight Watchers Smart Points values. The categories utilized are: 

1. Low: less than 10
2. Medium: 10 to less than 20
3. High: 20 or greater

"if_else" statements are used to create the new variable based on this categorization.

## `search_by_2_nutrs`

The third function is called `search_by_2_nutrs`, and it allows the user to search by two nutrients and select any range for two nutrients. In this function, the initial nutrients search based on two of choosing among number of calories, grams of carbohydrates, grams of protein, and grams of fat. And the user also can decide how many recipes would like to get. For example, if the user likes fitness, this function can use to search the nutrients with low calories but high protein. The full code of the function is as follows:  

```{r}
search_by_2_nutrs <- function(first_nutrient, first_from, first_to, sec_nutrient, sec_from, sec_to, number){
  base<-"https://api.spoonacular.com/recipes/findByNutrients"
  call2 <- paste(base,"?min",first_nutrient,"=",first_from,"&max",first_nutrient,"=", first_to,
                 "&min",sec_nutrient,"=",sec_from,"&max",sec_nutrient,"=", sec_to,
                 "&number=",number,"&apiKey=90c7af330c5f45a4a1b709068daed452",sep="")
  get_2_nutrs <- GET(call2)
  get_2_nutrs_text <- content(get_2_nutrs, "text")
  get_2_nutrs_json <- fromJSON(get_2_nutrs_text, flatten = TRUE)
  get_2_nutrs_tb <- as_tibble(get_2_nutrs_json) %>%
    select("id","title","calories","protein","fat","carbs")
  return(get_2_nutrs_tb)
}
```

# Exploratory Data Analysis
Now that we have set up our functions and run some test searches, we can do some exploratory data analysis of the results.First, let's look at some contingency tables and bar plots of Weight Watcher points categories, grouped by vegetarian to see if vegetarian results tend to be lower point values.

```{r}
#Create contingency table from calorie test results
table(calorie_test$WW_category, calorie_test$vegetarian, dnn=c("WW Category", "vegetarian"))
#Create contingency table for ingredient test results
table(ingredient_test$WW_category, ingredient_test$vegetarian, dnn=c("WW Category", "vegetarian"))
#Plot my tests as bar plots of Weight Watchers category grouped by vegetarian.
g <- ggplot(calorie_test, aes(x=WW_category))
g + geom_bar(aes(fill=vegetarian), position="dodge") +
  labs(title="Bar plot of WW Points for 100-500 Calorie Search", x="WW Smart Points Category")
  
g <- ggplot(ingredient_test, aes(x=WW_category))
g + geom_bar(aes(fill=vegetarian), position="dodge") +
  labs(title="Bar plot of WW Points for Ingredient=Cherry", x="WW Smart Points Category")
```
  
The most noticeable result here is that we get many more vegetarian results with our fruit ingredient choice vs. a calorie-based search.

Next, let's see if anything seems to trend with aggregate likes. Possibly results with higher calories tend to get more likes since they would tend to have more sugar or fat to make them taste good. Possibly higher health scores tend to get more likes because of their perceived health halo.  
```{r}
#Create scatterplots with aggregate likes as the response variable.
g <- ggplot(calorie_test, aes(x=calories, y=aggregateLikes))
g + geom_point() +
  labs(title="Aggregate Likes vs Calories from 100-500 Calorie Search Results", x="Calories", y="Aggregate Likes")
g <- ggplot(ingredient_test, aes(x=healthScore, y=aggregateLikes))
g + geom_point() + 
  labs(title="Aggregate Likes vs Calories from Cherry Ingredient Search Results", x="Health Score", y="Aggregate Likes")
```
  
The most notable result in these plots is that not very many of these recipes get very many likes. There are a few clear outliers with many more likes than the rest.  

We are now interested in looking at whether there is a difference in the spread of health score results for vegetarian and non-vegetarian recipes utilizing some box plots.  

```{r}
#Next box plot of calories or health score broken out by vegetarian and not
g <- ggplot(calorie_test, aes(x=vegetarian, y=calories))
g + geom_boxplot()+
  labs(title =  "Boxplot of calories for 100-500 calorie search")
g <- ggplot(ingredient_test, aes(x=vegetarian, y=healthScore))
g + geom_boxplot() + 
  labs(title =  "Boxplot of health scores for cherry ingredient search")
```
  
The plots show a smaller spread of the calorie results for vegetarian recipes and some high outliers in the health scores for some of the vegetarian results with cherry as an ingredient.  

One more thing that we could look at are the histogram distributions of calories for the calorie search and health score for the ingredient search results to see if any interesting distributions arise.  

```{r}
#Now, here are some histograms
g <- ggplot(calorie_test, aes(x=calories))
g + geom_histogram(bins=10, color = "black", fill = "lightblue",
size = 1) + 
  labs(title="Histogram of calories for recipes from 100-500 calories")
g <- ggplot(ingredient_test, aes(x=healthScore))
g + geom_histogram(bins=10, color = "black", fill = "lightblue",
size = 1) + 
  labs(title="Histogram of health scores for recipes with cherry as an ingredient")
```
  
The histograms show the results are skewed to the right in both distributions with peaks at relatively low numbers of calories and health scores.

Moreover, we can search any two nutrients from calories, protein, fat, carbohydrates, and see the recipes what we would like to choose. Let us take a look if we like fitness, and hope to choose food with low calories but high protein. We can give the calorie range and protein range. The full code using `search_by_2_nutrs` is as follows:

```{r}
cal_prot_test <- search_by_2_nutrs(first_nutrient="Calories", first_from=0, first_to=1500,
                                  sec_nutrient="Protein", sec_from=30, sec_to=100, number=100)

cal_prot_test$protein <- as.numeric(substr(cal_prot_test$protein,1, nchar(cal_prot_test$protein)-1))
cal_prot_test$fat <- as.numeric(substr(cal_prot_test$fat,1, nchar(cal_prot_test$fat)-1))
cal_prot_test$carbs <- as.numeric(substr(cal_prot_test$carbs,1, nchar(cal_prot_test$carbs)-1))
cal_prot_test
```

Let's use scatter plots to see the relationship between calories and fat, calories and carbohydrates.

```{r}
g <- ggplot(cal_prot_test, aes(x = fat, y = calories))
g + geom_point() + geom_smooth(method = lm, col = "Red") +
  labs(title = "Scatter plot of calories vs fat")

g <- ggplot(cal_prot_test, aes(x = carbs, y = calories))
g + geom_point() + geom_smooth(method = lm, col = "Red") +
  labs(title = "Scatter plot of calories vs carbohydrates")
```

According to the plots above, we can see that the fat has a stronger slope for calories than carbohydrates. Therefore, fat has more effect on calories than carbohydrates.

We can continue to evaluate calories, fat, carbohydrates to see if there are interesting finds. We can compare fat to carbohydrates, in order to see how to choose food with fat and carbohydrates. Firstly, add a new output variable `calories_level`. This variable did not exist in the source data and had to be created using the values of calories. The categories utilized are: 

1. Low: less than 400
2. Medium: 400 to less than 800
3. High: greater than 800

```{r}
cal_prot_test <- cal_prot_test %>% 
  arrange(calories) %>% 
  mutate(calorie_level = ifelse(cal_prot_test$calories >= 800, "3",
                              ifelse(cal_prot_test$calories >= 400, "2",
                                     ifelse(cal_prot_test$calories >= 0, "1"))),
         calorie_level = factor(calorie_level,
                              level = c(1,2,3),
                              labels = c("low","medium","high")))

cal_prot_test <- cal_prot_test %>% select(calorie_level, calories, protein, fat, carbs, title)
cal_prot_test
```
Secondly, add other two output variables `fat_level` and `carbs_level`. Both `fat_level` and `carbs_level` categories utilized are: 

1. Low: less than 40
2. Medium: 40 to less than 80
3. High: greater than 80

```{r}
cal_prot_test <- cal_prot_test %>% arrange(fat) %>% 
  mutate(fat_level = ifelse(cal_prot_test$fat >= 80, "3",
                            ifelse(cal_prot_test$fat >= 40,"2",
                                   ifelse(cal_prot_test$fat >= 0, "1"))),
         fat_level = factor(fat_level,
                            level = c(1,2,3),
                            labels = c("low","medium","high")))

cal_prot_test <- cal_prot_test %>% arrange(carbs) %>%          
  mutate(carb_level = ifelse(cal_prot_test$carbs >= 80, "3",
                              ifelse(cal_prot_test$carbs >= 40, "2",
                                     ifelse(cal_prot_test$carbs >= 0, "1"))),
         carb_level = factor(carb_level,
                              level = c(1,2,3),
                              labels = c("low","medium","high")))

cal_prot_test <- cal_prot_test %>% 
  select(calorie_level, calories, fat_level, fat, carb_level, carbs, protein, title)
cal_prot_test
```
Now, let's use the contigency tables and bar plots to check the results.

```{r}
table(cal_prot_test$calorie_level, cal_prot_test$fat_level, dnn=c("calorie_level", "fat_level"))
table(cal_prot_test$calorie_level, cal_prot_test$carb_level, dnn=c("calorie_level", "carb_level"))
```

Through the tables, we can conclude that medium calorie level with low fat level have the most recipes for choosing calories and fat, and medium calorie level with low carbohydrate level have the most recipes for choosing calories and carbohydrate. In total, medium calorie level with low fat level have the most recipes, which is 46 counts.

```{r}
g <- ggplot(cal_prot_test, aes(x = fat_level))
g + geom_bar(aes(fill = calorie_level), position = "dodge") +
  labs(title = "Bar plot of calorie_level vs fat_level")
g <- ggplot(cal_prot_test, aes(x = carb_level))
g + geom_bar(aes(fill = calorie_level), position = "dodge") +
  labs(title = "Bar plot of calorie_level vs carb_level")
```

From bar plots above, we can find that for calorie level vs fat level, as fat increasing, all counts for three calorie levels are decreasing. For calorie level vs carbohydrate level, as carbohydrates increasing, the counts for medium and high calorie levels are decreasing, but counts for low calorie level is increasing from low carbohydrate level to medium carbohydrate level then decreasing from medium carbohydrate level to high carbohydrate level.

# Conclusion
We hope that this vignette has given you some insight into working with APIs. In summary, what we did throughout this vignette was build some functions to allow for some flexible user defined searches, ran a few example searches, and did some exploratory data analysis with the results of those searches. Once the functions are built, there is nearly endless data that can be explored. Happy searching!
